BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications

### 2017-03-29 Paper2/BOAST.tex
Modified affiliations.
Rewrote workflow description. Answers point 1) from reviewer.

%% or include affiliations in footnotes:
\author{
   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{6}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{2}Centre National de la Recherche Scientifique}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Grenoble Alpes}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/SP2M/L\_Sim}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{6}Aix Marseille Univ., CNRS, Centrale Marseille, LMA, Marseille, France.}
}

@@

Figure~\ref{fig:boast_workflow} illustrates the envisioned workflow and program
structure. The user starts from an application kernel (either designed or
implemented), and writes it in a dedicated language (Step 1). The language
provides enough flexibility for the kernel to be meta-programmed with several
orthogonal optimizations. Then the developer selects a kernel configuration and
a target language. Those parameters define the output source code that will be
generated by BOAST (Step 2). The resulting code source is then built according
to the user specified compiler and options (Step 3). If input data are available
then the kernel can be benchmarked and tested for non regression. Based on the
results, other optimizations can be selected (Step 4) or new optimizations can
be added to the BOAST sources. The process can be repeated until a good
candidate is found on the target platform. The resulting kernel is then added to
the program (Step 5).

Several steps in the workflow can be automated: optimization and compiler flags
exploration, non regression testing, benchmarking, \ldots This automation can be
scripted by the user or by interfacing other dedicated tools. In order
to achieve those goals three aspects should be considered: code description,
code generation, and kernel execution run-time.

>>>>

%% or include affiliations in footnotes:
\author{
   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{6}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{2}Centre National de la Recherche Scientifique}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Grenoble Alpes}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/MEM/L\_Sim}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}\\
   \IEEEauthorblockA{\IEEEauthorrefmark{6}Aix Marseille Univ., CNRS, Centrale Marseille, LMA, Marseille, France.}
}

@@

Figure~\ref{fig:boast_workflow} illustrates the workflow and program structure.
The user starts from an application kernel (either designed or implemented),
and writes it in a dedicated language (Step 1). The language provides enough
flexibility for the kernel to be meta-programmed with several orthogonal
optimizations. From this set of optimizations, possible languages targets, and
compilation options the user can design an optimization space to explore. This
optimization space can contains rules to remove infeasible candidates. BOAST
provides the mechanisms to specify those optimization spaces and enforce the
users rules.

Once this optimization space is designed, the user selects an optimization
strategy, brute force and genetic algorithm are provided by BOAST, or can
design his own. BOAST will then evaluate the different candidates generated by
the optimization program. The candidate's parameters define the output source
code that will be generated by BOAST (Step 2). The resulting code source is
then built according to the specified compiler and options (Step 3). The kernel
can be benchmarked and tested for non regression. Based on the results, other
optimizations can be selected (Step 4).The process can be repeated until a good
candidate is found on the target platform. The resulting kernel is then added
to the program (Step 5).

Of course, this workflow if not the only possible use of BOAST.  Especially,
during the kernel description phase the user can use the BOAST framework to
test his existing optimizations and analyse the results in order to design new
optimizations. He can possibly do it using external tools.

In order too achieve those results three aspects should be considered: code
description, code generation, and kernel execution run-time.
